---
title: "Project for Machine Learning"
author: "Xiang Jiang"
date: "August 23, 2014"
output: html_document
---

### Preprocessing

Load data and use medianImputation to repalce missing values.
```{r,message=FALSE}
library(caret)

# Load data
train <- read.csv('/Users/xiangjiang/Documents/git/Data Science at Coursera/8.Practical Machine Learning/pml-training.csv')

test <- read.csv('/Users/xiangjiang/Documents/git/Data Science at Coursera/8.Practical Machine Learning/pml-testing.csv')

train <- train[,-1]
test <- test[,-1]

# remove columns with too much NA
col2remove <- apply(!is.na(train), 2, sum)>5000
train <- train[,col2remove]
test <- test[,col2remove]

train$user_name <- factor(train$user_name)
train$cvtd_timestamp <-  factor(train$cvtd_timestamp)
train$new_window <- factor(train$new_window)
train$classe <- factor(train$classe)

test$user_name <- factor(test$user_name)
test$cvtd_timestamp <-  factor(test$cvtd_timestamp)
test$new_window <- factor(test$new_window)


# get numeric columns
numeric_cols <- sapply(train, is.numeric)

# preprocessing for training data
prep <- preProcess(train[, numeric_cols], method = c("center","scale","medianImpute"))
newTrain <- predict(prep, train[, numeric_cols])
newTrain <- cbind(user_name=train$user_name,cvtd_timestamp=train$cvtd_timestamp,new_window=train$new_window,newTrain)

# preprocessing for testing data
newTest <- predict(prep, test[, numeric_cols])
newTest <- cbind(user_name=test$user_name,cvtd_timestamp=test$cvtd_timestamp,new_window=test$new_window,newTest)

```

### Training the Model

The model is trained using decision tree and cross-validation is used to select the best model.

```{r,message=FALSE}
ctrl <- trainControl(method = "cv")

fit <- train(train[,]$classe~., method="ctree",data=newTrain[,], trControl = ctrl)
```

### Model Evaluation

The accuracy for acoss-validation are:
```{r,message=FALSE}
fit$results
```
Accuracy was used to select the optimal model using  the largest value.
The final value used for the model was mincriterion = 0.5, and classification accuracy was 97.7%.

### Making Predictions

The predictions for the test set are:
```{r,message=FALSE}
prediction <- predict(fit, newTest)
prediction
```

